Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-05 eurocg/paper.tex
beta 4.2

The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $. Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.  Also it is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.

\martin{why do we need the reference here? Maybe better put it when it's actually used.}

@@

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms.

@@

Computing the intersection graph requires to identify all the intersections in $ \Gamma $. Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings. Building the graph then takes $ O(|S|) $ time. When computing the intersection we can compute the stub lengths and their order as well.

\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}

@@

Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ k' $ has a parent $ p(u) \in V $. For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{k'}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e. the next longer stub length, intersects $ \ell_0(u) $. We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $. See Figure~\ref{fig:two_solutions} for an illustration. More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
\begin{align*}
	\solshort(u) = \max\{T_1(u),\dots,T_j(u)\} \\
	\sollong(u) = \max\{T_0(u),\dots,T_{k'}(u)\}
\end{align*}

\martin{-- maybe $\delta$ instead of $k'$\\
-- in the else case of $T_i$ long should be enough as always long $\ge$ short}

It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:

\begin{align}
	\label{rec:tree}
	T_i(u) = \ell_i(u) + \sum_{v \in c(u)}
	\begin{cases}
		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\
		\max\{\solshort(v),\sollong(v)\} & \text{else}
	\end{cases}
\end{align}

Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.

\begin{lemma}
	\label{lem:tree_correctness}	
	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
\end{lemma}

\martin{could also use long(r) in the lemma}

\begin{proof}
	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ t = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is the maximum value for some $ T_i(r) $'s and $ r $ the root. Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $. This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition. Further we know the stubs can be categorized into two sets relative to their parent. Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort(v) $ and $ \sollong(v) $ values for all $ v \in c(u) $, such that no two stubs intersect. Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
\end{proof}

\martin{could also use induction on the tree structure. not sure what's better.}

\begin{figure}
	\centering
	\includegraphics{speedup}
	\caption{The situation in Lemma~\ref{lem:compute_vertex}.}
	\label{fig:speedup}
\end{figure}
Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in cubic time depending on the number of vertices in $ C $. Using the order on the stub lengths though we  can improve this to a quadratic running time by solving $ T_i(u) $ in linear time for one $ u \in V $.

\martin{the naïve solution would take $O(m\cdot k^2)$ and the improvement $O(mk)$, right?}

\begin{lemma}
	\label{lem:compute_vertex}
	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{k'} $ for every $ u \in V $ with degree $ k' $. The set $ \{T_i(u) \mid 0 \leq i < k' \} $ can be computed in $ O(k') $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}

\martin{The lemma should state the time to compute not only the $T_i(u)$ values but also long and short}

\begin{proof}
	Let $ u \in V $ be a vertex with degree $ k' < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children. The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order. First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children and the parent with their full segments. 
	\martin{we don't draw the parent yet; it's not included in $T_i(u)$. Moreover the children don't need to be all full segments, but they are unrestricted.}	
	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(k') $. The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $. Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $. Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:speedup}). By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order. Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $. Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
\end{proof}

>>>>

The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $. Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.

@@

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. Since every planar graph is representable as intersection graph of line segments~\cite{chalopin2009every}. we can also encounter every tree or cactus. In both cases we present polynomial-time dynamic programming algorithms.

@@

%Computing the intersection graph requires to identify all the intersections in $ \Gamma $. Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings. Building the graph then takes $ O(|S|) $ time. When computing the intersection we can compute the stub lengths and their order as well.
%
%\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
%\fabian{add to prelims}

@@

Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $. For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{\delta}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e. the next longer stub length, intersects $ \ell_0(u) $. We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $. See Figure~\ref{fig:two_solutions} for an illustration. More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
\begin{align*}
	\solshort(u) = \max\{T_1(u),\dots,T_j(u)\} \\
	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}
\end{align*}
It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
\begin{align}
	\label{rec:tree}
	T_i(u) = \ell_i(u) + \sum_{v \in c(u)}
	\begin{cases}
		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\
		\sollong(v) & \text{else}
	\end{cases}
\end{align}
Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.

\begin{lemma}
	\label{lem:tree_correctness}	
	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
\end{lemma}

\begin{proof}
	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $. Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $. Now let $ u \in V $ with degree $ \delta $ be an inner vertex. The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. $ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in. Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.

%	and $ t = \sollong(r) $ is the maximum value for some $ T_i(r) $'s and $ r $ the root. Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $. This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition. Further we know the stubs can be categorized into two sets relative to their parent. Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort(v) $ and $ \sollong(v) $ values for all $ v \in c(u) $, such that no two stubs intersect. Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
\end{proof}

\begin{figure}
	\centering
	\includegraphics{speedup}
	\caption{The situation in Lemma~\ref{lem:compute_vertex}.}
	\label{fig:speedup}
\end{figure}
Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in $ O(mk^2) $. Using the order on the stub lengths though we can improve this to $ O(mk) $ by solving $ T_i(u) $ in linear time for one $ u \in V $.

\begin{lemma}
	\label{lem:compute_vertex}
	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $. We can compute $ \solshort(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children. The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order. First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children with their full segments. 
	\fabian{make clear that we do not draw them, but could potentially do so if one wished for it}	
	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(\delta) $. The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $. Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $. Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:speedup}). By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order. Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $. Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.

	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most. This can be done in $ O(\delta) $ by simply sweeping over the values.
\end{proof}
